---
title: 'PHW251 Fall 2022 Project Milestone #2'
author: "Aubrey Robinson, Elizabeth Guzman, Tin Ho (Group Z)"
date: "Due: 2022-10-03"
output: pdf_document
---


# Task for Milestone 2

[Info per bCourse](https://bcourses.berkeley.edu/courses/1517565/assignments/8454501)


- Description of dataset
  - What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
  - How does the dataset relate to the group problem statement and question?
- Import statement
  - NOTE: Please use datasets available in the PHW251 Project Data github repo (Links to an external site.) (this is important to make sure everyone is using the same datasets)
  - Use appropriate import function and package based on the type of file
  - Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)
  - Document the import process
- Identify data types for 5+ data elements/columns/variables
  - Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
  - Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
  - Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type?
- Provide a basic description of the 5+ data elements
  - Numeric: mean, median, range
  - Character:  unique values/categories
  - Or any other descriptives that will be useful to the analysis



# Loading R libraries

```{r load-lib, include=TRUE, echo=TRUE}
library(tidyverse)
library(dplyr)

```

# Loading data

```{r load-data, echo=T, results="markup"}

demographics_path = 'data/ca_county_demographics.csv'
demographics_data = read_csv( demographics_path )

summary( demographics_data )

# more R code tbd.


```



\newpage

# Answers to Question 1

tbd

```{r Q1}

# r code here tbd ...
```
\newpage

# Answers to Question 2

tbd ...

