---
title:    '~~~ DRAFT DRAFT DRAFT ~~~'
subtitle: 'PHW251 Fall 2022 Project Milestone #2'
author:   "Aubrey Robinson, Elizabeth Guzman, Tin Ho (Group Z)"
date:     "Due: 2022-10-03"
output: pdf_document
---

# ** DRAFT  DRAFT **

# Task for Milestone 2

[Info from bCourse](https://bcourses.berkeley.edu/courses/1517565/assignments/8454501)
re-listed below as separate sections.


# Loading R libraries

```{r load-lib, include=TRUE, echo=TRUE}
library(tidyverse)
library(dplyr)

```
\newpage


# Description of dataset

-  What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
We have 3 different data sets that we will eventually be merging into a 
single one. Thye are county, mortality, and HCAI funding data. 
The county demographic data is from the 2012 California census. The mortality
data was sourced from the CA Dept of Public Health's Death 
Profiles by County from 2014-2020. The HCAI funding data set comes from
the CA.gov website which is the California open data portal, which includes
data from 2013-2022.

- How does the dataset relate to the group problem statement and question?

The demographic dataset will help us identify counties that meet OHE's 
criteria of low population per square mile, high median age, and a high
proportion of renters vs. homeowners.
The mortality data will help the team identify 
counties with high occurrences of death caused by chronic health conditions. 
The HCAI funding dataset will help us identify which projects are in closure
and which counties have been receiving a low amount of funding to help us 
determine which counties should receive more funding.


\newpage



# Import statement
  - NOTE: Please use datasets available in the PHW251 Project Data github repo (Links to an external site.) (this is important to make sure everyone is using the same datasets)
  - Use appropriate import function and package based on the type of file
  - Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)
  - Document the import process



## Loading data

We will use dplyr's read_csv to import the data, 
since they are all presented as csv files.


```{r load-data, echo=T, results="markup"}

#   - Utilize function arguments to control relevant components 
#     (i.e. change column types, column names, missing values, etc.)

demographics_path = 'data/ca_county_demographics.csv'
demographics_data = read_csv( demographics_path, 
                              na = c("", "NA", "-"), 
                              show_col_types=F )
# the first column contains id number, but it is unamed, so renaming it
# rest of the columns have reasonable names, so left them as is.
demographics_data = rename( demographics_data, id="...1")

# View(demographics_data)

# Loading Mortality Data 

mortality_path <- 'data/ca_county_mortality.csv'
mortality_data <- read_csv(mortality_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F)

mortality_data <- mortality_data %>% mutate_all(~replace( ., is.na(.), 0))

# View(mortality_data)
# There are NA values in mortality_data, so we need to replace NA w/ 0
 

# Loading HCAI funding Data 

funding_path = 'data/hcai_healthcare_construction.csv'
funding_data <- read_csv(  funding_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F       )

# finding where in the data frame there is an 'na'
# https://www.geeksforgeeks.org/find-columns-and-rows-with-na-in-r-dataframe/
funding_data_no_CtyColl = select(funding_data, -c("Collection of Counties"))
which(is.na(funding_data_no_CtyColl), arr.ind=T)
# and we find that only the column "Collection of Counties" has 'na'
# we will leave this for now since it may just be a colloquial reference
# unimportant for our data analysis.
# no replacement for na with 0 will be done on this data frame.


# the Costs column has human data, eg $50,890,315.00
# and we need to strip the dollar sign, the commas, 
# and convert them to numbers.  
# ref: https://stackoverflow.com/questions/31944103/convert-currency-with-commas-into-numeric
# we create a new column for this called "Numeric_Cost", 
# but could have potentially done an in-place replacement
funding_data = funding_data %>% 
  mutate(Numeric_Cost = as.numeric(
    gsub( '[$,]', '', funding_data[["Total Costs of OSHPD Projects"]] ) 
  ))

```



\newpage


# Identify data types for 5+ data elements/columns/variables
  
  
  - Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
  
  - Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
  
  - Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type?
  

```{r data-types}


str( demographics_data )
typeof( demographics_data[["name"]])
typeof( demographics_data[["pop2012"]])

str(mortality_data)
typeof(mortality_data[["County"]])
typeof(mortality_data[["Geography_Type"]])
typeof(mortality_data[["Cause"]])
typeof(mortality_data[["Count"]])

str(funding_data)
typeof(funding_data[["Total Costs of OSHPD Projects"]])
typeof(funding_data[["Numeric_Cost"]])
```

## demographics_data

- The name column holds a variable of character string type, and seems to contain the name of counties.  We may consider converting this into a Factor, will do so later on if we find such conversion to be useful.  
- pop2012 is a numeric field containing the number of people of the named county, in 2012.  We can perform computation such as mean calculaations on this field, see below, so there isn't likely any need for conversion.

## mortality_data

- County is a character string. 
- Geography_Type is a character string. 
- Cause is a character string.
- Count is a number data type. It is the count of events.

## HCAI funding data

- County is a character string.  However, there is also a number in it.  
eg "01 - Alameda".  To join this data frame with the others, 
there is likely some manipulation needed to strip out the number part
eg remove "01 - " and leave it with county names only
- Total Costs of OSHPD Projects was meant to be a numeric field,
but it has dollar sign and commas, and so a string parsing to strip them out,
and converted to numeric value, was done during the csv data import process above.
  
  
\newpage 

# Provide a basic description of the 5+ data elements

  - Numeric: mean, median, range
  - Character:  unique values/categories
  - Or any other descriptives that will be useful to the analysis



## demographics_data

Code to count number of unique counties:

```{r count-uniq-counties }

# Python style prinf() function per 
# https://stackoverflow.com/questions/13023274/how-to-do-printf-in-r
printf <- function(...) cat(sprintf(...))


# count number of unique name (ie counties)
uniq_counties = unique( demographics_data[["name"]]) %>% as.data.frame()
uniq_counties_count = count(uniq_counties)

printf( "The number of unique counties in the demographics data set was: %g", 
        uniq_counties_count )
```

Code to find statistics of numerical data (population in 2012):

```{r data-stats}

summary( demographics_data )

# IQR for numeric data
Q1     = quantile(  demographics_data[["pop2012"]], probs = 0.25, na.rm=T )
Median = median(    demographics_data[["pop2012"]],               na.rm=T )
Mean   = mean(      demographics_data[["pop2012"]],               na.rm=T )
Q3     = quantile(  demographics_data[["pop2012"]], probs = 0.75, na.rm=T )
Q1n    = round( Q1[[1]], 2 )
Q3n    = round( Q3[[1]], 2 )


printf( "The Mean for pop2012 in the demographics data set was found to be: %g", 
        Mean )
printf( "The interquartile range for pop2012 set was found to range from %g to %g",
        Q1n, Q3n )

```

- For the name column, it does not make much sense to talk about means or range, but we did found that our data set has 58 unique counties (ie, all counties of California is present in this data set)
- pop2012 has a mean of 650129, and an inter-quartile range of (48492, 645995)

## Mortality_data Descriptions

```{r }

#Summary of full dataset

summary(mortality_data)

# Unique characters: County, Geography Type, Causes

mortality_counties <- unique(mortality_data$County)
mortality_counties

mortality_geo_type <- unique(mortality_data$Geography_Type)
mortality_geo_type

mortality_causes <- unique(mortality_data$Cause)
mortality_causes

# Summary for Numeric Data - Count 

summary(mortality_data$Count)

```

- There are 58 unique characters in the variable "County". 
- There are 2 unique characters in the varaible "Geography_Type". 
If Geography_Type is Resident, the death was of a CA resident but may have 
occurred out of state. If Geography_Type is Occurrence, the death occurred in
CA but the person may not have been a CA resident. Our research question asks 
us to focus on "occurrences".
- There are 15 unique "Causes" of death. We are interested in
chronic health conditions. 
- The mean Count of mortality events is 189.8. The median is 0. and the
IQR is 41.


## HCAI Funding Data Descriptions


```{r funding-stats}

summary(funding_data)

```

- we find that the OSHPD Project costs has a mean value of $59M and a median of $2.8M
