---
title: 'PHW251 Fall 2022 Project Milestone #2'
author: "Aubrey Robinson, Elizabeth Guzman, Tin Ho (Group Z)"
date: "Due: 2022-10-03"
output: pdf_document
---


# Task for Milestone 2

[Info from bCourse](https://bcourses.berkeley.edu/courses/1517565/assignments/8454501)
re-listed below as separate sections.


# Loading R libraries

```{r load-lib, include=TRUE, echo=TRUE}
library(tidyverse)
library(dplyr)

```
\newpage


# Description of dataset

-  What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)

- How does the dataset relate to the group problem statement and question?


(Answeres tbd)


\newpage



# Import statement
  - NOTE: Please use datasets available in the PHW251 Project Data github repo (Links to an external site.) (this is important to make sure everyone is using the same datasets)
  - Use appropriate import function and package based on the type of file
  - Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)
  - Document the import process




## Loading data

```{r load-data, echo=T, results="markup"}

#   - Utilize function arguments to control relevant components 
#     (i.e. change column types, column names, missing values, etc.)

demographics_path = 'data/ca_county_demographics.csv'
demographics_data = read_csv( demographics_path, 
                              na = c("", "NA", "-"), 
                              show_col_types=F )
# the first column contains id number, but it is unamed, so renaming it
# rest of the columsn have reasonable names, so left them as is.
demographics_data = rename( demographics_data, id="...1")




# more R code tbd.




```



\newpage


# Identify data types for 5+ data elements/columns/variables
  
  
  - Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
  
  - Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
  
  - Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type?
  

```{r data-types}

# r code here tbd on mean, median, range ...

str( demographics_data )
typeof( demographics_data[["name"]])
typeof( demographics_data[["pop2012"]])


```

## demographics_data

- The name column holds a variable of character string type, and seems to contain the name of counties.  We may consider converting this into a Factor, will do so later on if we find such conversion to be useful.  
- pop2012 is a numeric field containing the number of people of the named county, in 2012.  We can perform computation such as mean calculaations on this field, see below, so there isn't likely any need for conversion.


(more tbd)
  



  
\newpage 

# Provide a basic description of the 5+ data elements

  - Numeric: mean, median, range
  - Character:  unique values/categories
  - Or any other descriptives that will be useful to the analysis

```{r data-stats}

# r code here tbd on mean, median, range ...

summary( demographics_data )
```


## demographics_data

Code to count number of unique counties

```{r count-uniq-counties }

# Python style prinf() function per 
# https://stackoverflow.com/questions/13023274/how-to-do-printf-in-r
printf <- function(...) cat(sprintf(...))


# count number of unique name (ie counties)
uniq_counties = unique( demographics_data[["name"]]) %>% as.data.frame()
uniq_counties_count = count(uniq_counties)

printf( "The number of unique counties in the demographics data set was: %g", 
        uniq_counties_count )


# IQR for numeric data
Q1     = quantile(  demographics_data[["pop2012"]], probs = 0.25, na.rm=T )
Median = median(    demographics_data[["pop2012"]],               na.rm=T )
Mean   = mean(      demographics_data[["pop2012"]],               na.rm=T )
Q3     = quantile(  demographics_data[["pop2012"]], probs = 0.75, na.rm=T )
Q1n    = round( Q1[[1]], 2 )
Q3n    = round( Q3[[1]], 2 )


printf( "The Mean for pop2012 in the demographics data set was found to be: %g", 
        Mean )
printf( "The interquartile range for pop2012 set was found to range from %g to %g",
        Q1n, Q3n )

```

- For the name column, it does not make much sense to talk about means or range, but we did found that our data set has 58 unique counties (ie, all counties of California is present in this data set)
- pop2012 has a mean of 650129, and an inter-quartile range of (48492, 645995)

