---
title: "Milestone 4 Group Z"
output: html_document
date: '2022-11-19'
author:   'Aubrey Robinson, Elizabeth Guzman, Tin Ho'
---

# Web Access to knitted HTML output

We have decided to take a leap and try plot_ly and knit to html output for this milestone.
The html page is hosted as [Github Pages](https://pages.github.com/) and is accessible
in this address:
[https://tin6150.github.io/phw251_group_z/milestone4_groupZ.html](https://tin6150.github.io/phw251_group_z/milestone4_groupZ.html)


# Project and Milestone overview

[{Pasted to group's gdoc}](https://docs.google.com/document/d/1u4-f0KCUB0QGxWI1VVgfEgEaTmQbt-Jk1uIgWP-0VYk/edit#)


# R setup and data preparation

(This code block omitted for brevity, please refer to source at our
[github repo](https://github.com/tin6150/phw251_group_z) )

```{r load-lib, include=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(kableExtra)
library(lubridate)
library(ggplot2)
library(plotly)

# omit warning messages from library load
# https://stackoverflow.com/questions/45399587/how-to-remove-warning-messages-in-r-markdown-document
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```


```{r massage mortality data, echo=FALSE, include=FALSE}

## Loading & Cleaning data : Mortality

#**Suggestions from Lauren: 
#  -confirm we're not double counting deaths in mortality data 
#     Yes we are filtering by Total Population
#  -summarize mortality rates by county and use rates instead of counts before joining 
#     Okay now calculated Rate as count / pop12 

# Loading Mortality Data 

mortality_path <- 'data/ca_county_mortality.csv'
mortality_data_raw <- read_csv(mortality_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F)

mortality_data <- mortality_data_raw %>% mutate_all(~replace( ., is.na(.), 0))

# View(mortality_data)
# There are NA values in mortality_data, so we need to replace NA w/ 0

mortality_data %>% select( Year ) %>% unique()
# Mortality data is from 2014 - 2020.

# subset mortality data for chronic conditions only

chronic_desc = c( 
  "Chronic lower respiratory diseases",
  "Diabetes mellitus",
  "Diseases of heart",
  "Essential hypertension and hypertensive renal disease",
  "Chronic liver disease and cirrhosis",
  "Alzheimer's disease",
  "Malignant neoplasms",
  "Nephritis, nephrotic syndrome and nephrosis",
  "Cerebrovascular diseases",
  "Parkinson's disease"
  )

chronic_mortality_data = mortality_data %>% 
  filter( Cause_Desc %in% chronic_desc )

num_rows_befor_filter = mortality_data %>% tally
num_rows_after_filter = chronic_mortality_data %>% tally
#num_rows_after_filter

num_rows_filtered = num_rows_befor_filter - num_rows_after_filter
num_rows_filtered


#View(chronic_mortality_data)

# chronic case count by county in the last 5 years
# Question prompt:
## Second, OHE’s director would also like to include a total count of mortality 
## from chronic health conditions 
## over the past few years into the county level analysis. 


chronic_by_county = chronic_mortality_data %>%
  filter(  Strata == "Total Population") %>%
  filter(  Geography_Type == "Occurrence" ) %>%
  select( -Geography_Type ) %>% # Geo type restricted to "Occurrence"
  filter( Year > 2016 )     %>%   # 2016-2020 = 5 years
  group_by( County ) %>%
  summarize( sum_count = sum(Count) )

```

```{r massage funding data, echo=FALSE, include=FALSE}

# Loading HCAI funding Data 

funding_path = 'data/hcai_healthcare_construction.csv'
funding_data <- read_csv(  funding_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F       )

CtyColl = select(funding_data, c("Collection of Counties")) %>%
  unique()

CtyColl 
## there are 4 collection of counties, 
## there aren't likely needed for data eval,
## but add lots of NA, so dropping them
#1 Bay Area Counties           
#2 NA                          
#3 Greater Sacramento Counties 
#4 Greater Los Angeles Counties

# finding where in the data frame there is an 'na'
# https://www.geeksforgeeks.org/find-columns-and-rows-with-na-in-r-dataframe/
## funding_data_no_CtyColl = select(funding_data, -c("Collection of Counties"))
funding_data = select(funding_data, -c("Collection of Counties"))

## which(is.na(funding_data_no_CtyColl), arr.ind=T)
# above returns 0 rows, ie
# we find that only the column "Collection of Counties" has 'na'
# we will leave this for now since it may just be a colloquial reference
# unimportant for our data analysis.
# no replacement for na with 0 will be done on this data frame.



# the Costs column has human data, eg $50,890,315.00
# and we need to strip the dollar sign, the commas, 
# and convert them to numbers.  
# ref: https://stackoverflow.com/questions/31944103/convert-currency-with-commas-into-numeric
# we create a new column for this called "Numeric_Cost", 
# but could have potentially done an in-place replacement
funding_data = funding_data %>% 
  mutate(Numeric_Cost = as.numeric(
    gsub( '[$,]', '', funding_data[["Total Costs of OSHPD Projects"]] ) 
  ))

funding_data = funding_data %>%
  mutate( county_name = str_sub( County, 6 )) %>%
  subset(select = -c(County)) %>%
  rename(County = county_name)

# subset HCAI funding data for those "In Closure" state

funding_data_closure <- funding_data %>%
  filter(`OSHPD Project Status` == "In Closure") %>%
  filter(`Data Generation Date` == as_date("2022-08-11")  )

  ## 2022-08-11 the latest avail data

    #group_by(County, Numeric_Cost, `OSHPD Project Status`) %>%  
  ## I (Tin) can no longer get the the gruop_by  to work... 



```



```{r spot checks, echo=FALSE, include=FALSE}


funding_data_closure %>% arrange( Numeric_Cost ) %>% head( 10 )

#View(funding_data)
#View(funding_data_closure)

funding_data %>% select( `Data Generation Date` ) %>% unique() %>% tally() ## 231 rows

selected_counties = c(
  "Inyo",
  "Siskiyou",
  "Mariposa",
  "Modoc",
  "Plumas"
)

funding_data_selected_counties = funding_data %>% 
  #filter( County == "Alameda" ) %>%
  #filter( County == "Inyo" ) %>%
  #filter( County %in% selected_counties  ) %>%
  filter( `OSHPD Project Status` == "In Closure" |
            `OSHPD Project Status` == "In Construction"    ) %>%
  #filter( `OSHPD Project Status` == "Pending Construction"  ) %>%
  #filter( `OSHPD Project Status` == "In Review"  ) %>%
  filter( `Data Generation Date` == 
            as_date( "2022-08-11" ) 
            #as_date( "2020-01-01" ) 
            #as_date( "2016-01-01" ) 
          )

chk1 = 
  ggplot( data = funding_data_selected_counties,
          aes( x = `Data Generation Date`, 
               y = Numeric_Cost             ) ) +
  geom_bar( stat="identity", position="dodge" ) +
  labs( title = 'Prj cost over time' ) + 
  facet_wrap( ~ County )


chk1


## some number may linger in closure for longer than other?
## so taking average of closure amount over say 5 months may lead to skewed data
## really just make sense to just use the last data point.


funding_data_allSources = funding_data %>% 
  #filter( County == "Alameda" ) %>%
  #filter( County == "Mariposa" ) %>%
  #filter( County == "Inyo" ) %>%
  filter( County %in% selected_counties  ) %>%
  filter( `Data Generation Date` > 
            #as_date( "2022-01-01" ) 
            #as_date( "2020-01-01" ) 
            as_date( "2016-01-01" ) 
          )


chk_allSrc = 
  ggplot( data = funding_data_allSources,
          aes( x = `Data Generation Date`, 
               y = Numeric_Cost             ) ) +
  geom_bar( stat="identity", position="dodge" ) +
  labs( title = '1 county, prj stage' ) + 
  facet_wrap( ~ `OSHPD Project Status` )


chk_allSrc


## there is likely more to the story about these costs
## expect to see approved project to have a peak in Review
## then the peak move to Pending Construction, In Construction
## then In Closure
## but maybe their cost is finance book remaining budget account?
## data doesn't fully make sense
## but it is just homework, take a snapshot of "In Closure", and let it be.

```


```{r massage demographics data, echo=FALSE, include=FALSE}


# Load demographics data

demographics_path = 'data/ca_county_demographics.csv'
demographics_data = read_csv( demographics_path, 
                              na = c("", "NA", "-"), 
                              show_col_types=F )

# the first column contains id number, but it is unamed, so renaming it
# rest of the columns have reasonable names, so left them as is.
demographics_data = rename( demographics_data, id="...1")

# View(demographics_data)


# massaging demographics data

demographics_data <- demographics_data %>%
  select(id, name, pop2012, pop12_sqmi, med_age, owner_occ, renter_occ) %>%
  rename( `County` = name )


## Determining Renters vs Homeowners Ratio
demographics_data = demographics_data %>%
  mutate( rent_own_ratio = renter_occ / owner_occ )


demographics_data <- 
  mutate(demographics_data, 
         rural_class=if_else( pop12_sqmi < 20, "rural", "not rural", missing=NULL))

# column to show whether county was rural or not
rural_not_rural <- demographics_data %>% 
  select( County, rural_class, pop12_sqmi, rent_own_ratio, med_age )  

rural_counties = rural_not_rural   %>%
  filter( rural_class == "rural" ) 
  #rename( `Class`  = rural_class )


#### consider a "low pop density" rather than a strict "rural"
#### --> not really good, 49 counties in this grouping

avg_pop_density = mean( demographics_data$pop12_sqmi )
demographics_data = demographics_data %>% 
  mutate( low_pop = if_else( pop12_sqmi < avg_pop_density , 
                                    TRUE,
                                    FALSE,
                                    NULL   ))


```


```{r demographics data complete set high rental, old age, echo=FALSE, include=FALSE}

high_rental_ratio = 
  rural_counties %>% arrange( desc( rent_own_ratio  ) ) 

high_med_age = 
  rural_counties %>% arrange( desc( med_age  ) ) 

## Determine any counties satisfy triple whammy rule

avg_rent_ratio = mean( demographics_data$rent_own_ratio )
avg_age        = mean( demographics_data$med_age  )

demographics_data = demographics_data %>% 
  mutate( high_med_age = if_else( med_age > avg_age , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) %>%
  mutate( high_rental = if_else( rent_own_ratio > avg_rent_ratio , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) 
  
triple_whammy = demographics_data %>%
  filter( rural_class  == "rural" &
          high_rental  == TRUE    &  
          high_med_age == TRUE       )

### so 0 county satisfy the triple whammy rule
### need to modify  selection criteria
```


```{r demographics data RURAL set only, high rental, old age, echo=FALSE, include=FALSE}

## Use only RURAL counties to calculate their averages, and what's above average

## Determine any RURAL counties satisfy double whammy rule
#View(rural_counties)

avg_rent_ratio_rural = mean( rural_counties$rent_own_ratio )
avg_age_rural        = mean( rural_counties$med_age  )



rural_demographics_data = demographics_data %>% 
  mutate( high_med_age_rural = if_else( med_age > avg_age_rural , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) %>%
  mutate( high_rental_rural = if_else( rent_own_ratio > avg_rent_ratio_rural , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) 

#-- c( avg_rent_ratio, avg_rent_ratio_rural, avg_age, avg_age_rural )
#-- sn50 =  rural_demographics_data %>% filter( rural_class == "rural" ) 
```

```{r tmp-double-whammy, include=FALSE, echo=FALSE}

rural_double_whammy = rural_demographics_data %>%
  filter( 
          high_rental_rural  == TRUE    &  
          high_med_age_rural == TRUE       )



rural_double_whammy  # only 3 counties in this list: Lake, Siskiyou, Inyo

rural_demographics_data %>% arrange( rent_own_ratio, med_age ) %>% head( 6 )

```


```{r dem relaxed criteria: old age, high rental in rural standard, echo=FALSE, include=FALSE}

## determine old age as > avg age for full data set
## then pick rural countries, 
## high rental in such rural counties.


#View(rural_counties)


focus_demographics_data = demographics_data %>% 
  mutate( high_rental_rural = if_else( rent_own_ratio > avg_rent_ratio_rural , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) %>%
  mutate( high_med_age = if_else( med_age > avg_age , 
                                    TRUE,
                                    FALSE,
                                    NULL   ))

#-- c( avg_rent_ratio, avg_rent_ratio_rural, avg_age, avg_age_rural )
#-- sn50 =  focus_demographics_data %>% filter( rural_class == "rural" ) 

```

```{r tmp-focus,  include=FALSE, echo=FALSE }

focus_dem = rural_demographics_data %>%
  filter( 
          high_rental_rural  == TRUE    &  
          high_med_age == TRUE       )    %>%
  select( County, rural_class, high_med_age, high_rental_rural )


rural_demographics_data %>%
  select( County, rural_class, med_age, high_med_age, high_rental_rural ) %>%
  arrange( high_rental_rural, desc(med_age) )

#View(focus_dem)  # now have 13 counties to choose from, but not all rural

```



```{r discussion-use-only,  include=FALSE, echo=FALSE}

## decide if need to output any of these tables
## likely need to refine further 

high_rental_ratio %>% head( 6 )

high_med_age %>% head( 6 )

rural_double_whammy %>% head( 6 )

focus_dem %>% head( 13 )

#It doesn't seems like we have high rental (ie > avg_rent) and old age (>avg_age)
#so we will have to judge subjectively
#-Tin


```



```{r calc chronic mortality rate, echo=F, include=F}
# calc chronic mortality rate using census pop12 as denominator

# First need to Join Chronics mortality data (by county, last 5 years)
# and demographics data

demographics_chronic = left_join( demographics_data, 
                                  chronic_by_county, 
                                  by = "County"      ) %>% 
  mutate( pct = ( sum_count / pop2012 ) * 100 )

# pct = "prevalence" percentage
```



```{r join tables, echo=FALSE, include=FALSE}


fund_dem_chron = inner_join( demographics_chronic,
                             funding_data_closure, 
                             by = "County" ) %>%
  mutate( fund_per_cap = Numeric_Cost / pop2012 )

viz_fund_dem_chron = fund_dem_chron %>%
  select( County, 
          pop12_sqmi,
          rural_class,  
          #low_pop,
          med_age,
          high_med_age,
          rent_own_ratio,
          high_rental,
          pct,                    # prevalence,         
          Numeric_Cost, 
          fund_per_cap,
          `Number of OSHPD Projects`,
          )

```


```{r draft-viz-for-discussion-only, echo=FALSE, include=FALSE }

#### for viz, ponder about using these, or derivative thereof 

feel_for_data = viz_fund_dem_chron %>% 
  filter( rural_class == "rural" ) 
  #filter( low_pop == TRUE )         # no good, 49 counties in this grouping

feel_for_data %>% 
  arrange( desc( fund_per_cap ) ) %>% head( 10 )
### above are all 0... which implies no fundings for them... should give them some love!?



```


\newpage 

# Visualizations & Code

Data needed to make decision:
-which counties share all 3 attributes: low pop(rural), high median age, 
& high proportion of renters vs homeowners
-which 5 counties have highest mortality rates due to chronic illness
-what counties have received little to no funding recently? ---not quite sure 
about this one

# Rural Counties Deserving More HCAI Funding

Table 1 shows that the 5 counties of Siskiyou, Inyo, Mariposa, Plumas and Modoc
have low population density (they qualify as Rural per National Rural Development Partnership's definition), 
have a high median age, and fairly high ratio of renters.
More importantly, they have one of the highest percentage of chronic diseases,
and no HCAI fundings that are "In Closure" as of 2022-08-11.

Note that the latest population data is from 2012.


```{r print quality table, fig.cap="fig_caption_fixme"}


viz_focus = viz_fund_dem_chron %>%
  filter( rural_class      == "rural",
          high_med_age     == TRUE 
          ##high_rental    == TRUE  
          )

focus_table = viz_focus %>%
  select( County, 
          pop12_sqmi,
          #rural_class,  
          #low_pop,
          med_age,
          #high_med_age,
          rent_own_ratio,
          #high_rental,
          pct,                    # prevalence,         
          Numeric_Cost, 
          #fund_per_cap,
          #`Number of OSHPD Projects`,
  ) %>% 
  arrange( desc( pct )) %>%
  rename(
    `Pop Density` = pop12_sqmi,
    `Median Age` = med_age,
    `Rent:Own Ratio` = rent_own_ratio,
    `% Chronic`      = pct,
    `HCAI Fund in 2022` = Numeric_Cost
  ) %>%  
  head( 5 ) 

kable( focus_table,
       booktabs=T,
       digits=c(0,1,1,2,2,0),
       format.args=list(big.mark=','),
       caption = "Rural Counties with high median age, rental ratio, and chronic disease rate",
       )

```

# Print Quality Plot or Chart

The following boxplot summarizes chronic disease mortality rates from all 
CA counties, grouped according to HCAI funding amounts for in closure projects 
as of August 2022. The funding amounts were categorized as "high" if they were 
above the mean amount, low if they were below the mean, and "no funding" if no 
funding for in closure projects was reported. 


(Audrey, so what is graph telling the audience?)

```{r}

funding_chronic <- funding_data %>% 
  filter( `OSHPD Project Status` == "In Closure") %>%
  filter( `Data Generation Date` == as_date( "2022-08-11")) %>%
  mutate(funding_amount = case_when(
    Numeric_Cost > 12239849 ~ "High Funding", 
    Numeric_Cost == 0 ~ "No Funding", 
    Numeric_Cost < 12239849 ~ "Low Funding"
  )) %>%
  inner_join(demographics_chronic, funding_data_all_counties, by = "County") %>%
  select(pct, County, funding_amount, rural_class, Numeric_Cost)

plot_ly(
  funding_chronic,
  y=~pct,
  color= ~funding_amount,
  type="box"
) %>%
   layout(
    title="Chronic Disease Mortality Rates & HCAI Funding",
    yaxis=list(title="Chronic Disease Rate"))
```

# Funding for select counties

```{r viz: funding for select counties}

counties_funding_comp = funding_data %>%
  filter( `OSHPD Project Status` == "In Closure"  ) %>%
  filter( `Data Generation Date` ==
            as_date( "2022-08-11" )
            #as_date( "2020-01-01" )
            #as_date( "2016-01-01" )
          ) %>%
  arrange( Numeric_Cost )

#fig2 = plot_ly( data = funding_data_1county ) %>%
#fig2 = plot_ly( data = funding_data_selected_counties ) %>%
fig2 = plot_ly( data = counties_funding_comp ) %>%
  add_trace(x = ~`County`,
            y = ~Numeric_Cost,
            type = 'bar',
            name = 'Funding across 58 counties',
            marker = list(color = 'rgb(187, 216, 228)'),
            hoverinfo = "text",
            text = ~paste(round(Numeric_Cost, 0), ' US$')  ) %>%
   layout(
            title="Funding across 58 counties"
         )

fig2


```


# Mortality Rate by County

Below is a graph of Mortality Rate for Chronic diseases (as defined by CDC)
across all 58 California counties.
The 5 counties of focus have Mortality rates of:
(TBD)


```{r viz: chronic mortality rates by counties}

chronic_focus_counties = demographics_chronic %>% 
  mutate( focusCounty = if_else(
    County %in% selected_counties,
    TRUE,
    FALSE
  ) ) %>%
  mutate( ctyColor = case_when(
    County == "Siskiyou" ~ "red", 
    County == "Inyo"     ~ "darkorange" , 
    County == "Mariposa" ~ "blue", 
    County == "Plumas"   ~ "darkblue",
    County == "Modoc"    ~ "darkcyan",
    TRUE                 ~ "darkyellow"
  ) )

fig3 = plot_ly( data = chronic_focus_counties ) %>%
  add_trace(x = ~`County`,
            y = ~pct,
            type = 'bar',
            #color = ~focusCounty,
            color = ~ctyColor,
            name = 'Chronic Mortality Rates by Counties',
            marker = list(color = 'rgb(187, 216, 228)'),
            hoverinfo = "text",
            text = ~paste('County: ', County, '<BR>Mortality Rate:', round(pct, 1))  ) %>%
   layout(
            title = "Chronic Mortality Rates by Counties", 
            yaxis = list( title="% Mortality Rate" )
         )

fig3

```
