---
title: "milestone4_groupZ"
output: pdf_document
date: '2022-11-19'
author:   'Aubrey Robinson, Elizabeth Guzman, Tin Ho (Group Z)'
---

# Project and Milestone overview

[Pasted to group's gdoc:](https://docs.google.com/document/d/1u4-f0KCUB0QGxWI1VVgfEgEaTmQbt-Jk1uIgWP-0VYk/edit#)


# R setup and load libraries

(this code block omitted for brevity)

```{r load-lib, include=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(kableExtra)

# omit warning messages from library load
# https://stackoverflow.com/questions/45399587/how-to-remove-warning-messages-in-r-markdown-document
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```

## Loading & Cleaning data from Milestone 4

**Suggestions from Lauren: 
  -confirm we're not double counting deaths in mortality data 
  -summarize mortality rates by county and use rates instead of counts before joining 

```{r load-data, echo=FALSE, include=FALSE}

# Load demographics data

demographics_path = 'data/ca_county_demographics.csv'
demographics_data = read_csv( demographics_path, 
                              na = c("", "NA", "-"), 
                              show_col_types=F )

# the first column contains id number, but it is unamed, so renaming it
# rest of the columns have reasonable names, so left them as is.
demographics_data = rename( demographics_data, id="...1")

# View(demographics_data)


# Loading Mortality Data 

mortality_path <- 'data/ca_county_mortality.csv'
mortality_data_raw <- read_csv(mortality_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F)

mortality_data <- mortality_data_raw %>% mutate_all(~replace( ., is.na(.), 0))

# View(mortality_data)
# There are NA values in mortality_data, so we need to replace NA w/ 0
 

# Loading HCAI funding Data 

funding_path = 'data/hcai_healthcare_construction.csv'
funding_data <- read_csv(  funding_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F       )

# finding where in the data frame there is an 'na'
# https://www.geeksforgeeks.org/find-columns-and-rows-with-na-in-r-dataframe/
funding_data_no_CtyColl = select(funding_data, -c("Collection of Counties"))

## which(is.na(funding_data_no_CtyColl), arr.ind=T)
# above returns 0 rows, ie
# we find that only the column "Collection of Counties" has 'na'
# we will leave this for now since it may just be a colloquial reference
# unimportant for our data analysis.
# no replacement for na with 0 will be done on this data frame.


# the Costs column has human data, eg $50,890,315.00
# and we need to strip the dollar sign, the commas, 
# and convert them to numbers.  
# ref: https://stackoverflow.com/questions/31944103/convert-currency-with-commas-into-numeric
# we create a new column for this called "Numeric_Cost", 
# but could have potentially done an in-place replacement
funding_data = funding_data %>% 
  mutate(Numeric_Cost = as.numeric(
    gsub( '[$,]', '', funding_data[["Total Costs of OSHPD Projects"]] ) 
  ))

```

```{r massage mortality data, echo=FALSE, include=FALSE}

# subset mortality data for chronic conditions only

chronic_desc = c( 
  "Chronic lower respiratory diseases",
  "Diabetes mellitus",
  "Diseases of heart",
  "Essential hypertension and hypertensive renal disease",
  "Chronic liver disease and cirrhosis",
  "Alzheimer's disease",
  "Malignant neoplasms",
  "Nephritis, nephrotic syndrome and nephrosis",
  "Cerebrovascular diseases",
  "Parkinson's disease"
  )

chronic_mortality_data = mortality_data %>% 
  filter( Cause_Desc %in% chronic_desc )

num_rows_befor_filter = mortality_data %>% tally
num_rows_after_filter = chronic_mortality_data %>% tally
#num_rows_after_filter

num_rows_filtered = num_rows_befor_filter - num_rows_after_filter
num_rows_filtered
```

```{r massage funding data, echo=FALSE, include=FALSE}

# subset HCAI funding data for those "In Closure" state

funding_data <- funding_data %>%
  group_by(County, Numeric_Cost, `OSHPD Project Status`) %>%
  filter(`OSHPD Project Status` == "In Closure") %>%
  filter(`Data Generation Date` == "2022-08-11")

funding_data = funding_data %>%
  mutate( county_name = str_sub( County, 6 )) %>%
  subset(select = -c(County)) %>%
  rename(County = county_name)

```


```{r massage demographics data, echo=FALSE, include=FALSE}

# massaging demographics data

demographics_data <- demographics_data %>%
  select(id, name, pop2012, pop12_sqmi, med_age, owner_occ, renter_occ)

## Determining Renters vs Homeowners Ratio
demographics_data = demographics_data %>%
  mutate( rent_own_ratio = renter_occ / owner_occ )



```

```{r join tables, echo=FALSE, include=FALSE}



```

\newpage 

#Visualizations

Data needed to make decision:
-which counties share all 3 attributes: low pop(rural), high median age, 
& high proportion of renters vs homeowners
-which 5 counties have highest mortality rates due to chronic illness
-what counties have received little to no funding recently? ---not quite sure 
about this one

#Print Quality Table

```{r}

```

#Print Quality Plot or Chart

```{r}

```

#Print Quality Table, Plot or Chart

```{r}

```

