---
title: "**DRAFT** Milestone 5 Group Z **DRAFT** "
output: html_document
date: '2022-11-23'
author:   'Aubrey Robinson, Elizabeth Guzman, Tin Ho'
---

**DRAFT**  **DRAFT**  **DRAFT** 

# Web access to knitted HTML output

We have decided to take a leap and try plot_ly and knit to html output for this milestone.
The html page is hosted as [Github Pages](https://pages.github.com/) and is accessible
in this address:
++TBD++ [https://tin6150.github.io/phw251_group_z/milestone5_groupZ.html](https://tin6150.github.io/phw251_group_z/milestone5_groupZ.html)


This is intended as final end user report, 
code blocks have been omitted for brevity, but it is available at
[our github repo](https://github.com/tin6150/phw251_group_z) )

```{r load-lib, include=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(kableExtra)
library(lubridate)
library(ggplot2)
library(plotly)
library(formattable)   # for currency()
library(DT)


# omit warning messages from library load
# https://stackoverflow.com/questions/45399587/how-to-remove-warning-messages-in-r-markdown-document
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```


```{r massage mortality data, echo=FALSE, include=FALSE}

## Loading & Cleaning data : Mortality

#**Suggestions from Lauren: 
#  -confirm we're not double counting deaths in mortality data 
#     Yes we are filtering by Total Population
#  -summarize mortality rates by county and use rates instead of counts before joining 
#     Okay now calculated Rate as count / pop12 

# Loading Mortality Data 

mortality_path <- 'data/ca_county_mortality.csv'
mortality_data_raw <- read_csv(mortality_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F)

mortality_data <- mortality_data_raw %>% mutate_all(~replace( ., is.na(.), 0))

# View(mortality_data)
# There are NA values in mortality_data, so we need to replace NA w/ 0

mortality_data %>% select( Year ) %>% unique()
# Mortality data is from 2014 - 2020.

mortality_data %>% select( "Cause_Desc" ) %>% unique() %>% tally()
# contain 15 diseases, but one is "All causes (total)"


# subset mortality data for chronic conditions only

chronic_desc = c( 
  "Chronic lower respiratory diseases",
  "Diabetes mellitus",
  "Diseases of heart",
  "Essential hypertension and hypertensive renal disease",
  "Chronic liver disease and cirrhosis",
  "Alzheimer's disease",
  "Malignant neoplasms",
  "Nephritis, nephrotic syndrome and nephrosis",
  "Cerebrovascular diseases",
  "Parkinson's disease"
  )

str( chronic_desc )

chronic_mortality_data = mortality_data %>% 
  filter( Cause_Desc %in% chronic_desc )

num_rows_befor_filter = mortality_data %>% tally
num_rows_after_filter = chronic_mortality_data %>% tally
#num_rows_after_filter

num_rows_filtered = num_rows_befor_filter - num_rows_after_filter
num_rows_filtered


#View(chronic_mortality_data)

# chronic case count by county in the last 5 years
# Question prompt:
## Second, OHE’s director would also like to include a total count of mortality 
## from chronic health conditions 
## over the past few years into the county level analysis. 


chronic_by_county = chronic_mortality_data %>%
  filter(  Strata == "Total Population") %>%
  filter(  Geography_Type == "Occurrence" ) %>%
  select( -Geography_Type ) %>% # Geo type restricted to "Occurrence"
  filter( Year >= 2016 )     %>%   # 2016-2020 = 5 years
  group_by( County ) %>%
  summarize( sum_count = sum(Count) )

```

```{r massage funding data, echo=FALSE, include=FALSE}

# Loading HCAI funding Data 

funding_path = 'data/hcai_healthcare_construction.csv'
funding_data <- read_csv(  funding_path, 
                           na= c("", "NA", "-"),
                           show_col_types=F       )

CtyColl = select(funding_data, c("Collection of Counties")) %>%
  unique()

CtyColl 
## there are 4 collection of counties, 
## there aren't likely needed for data eval,
## but add lots of NA, so dropping them
#1 Bay Area Counties           
#2 NA                          
#3 Greater Sacramento Counties 
#4 Greater Los Angeles Counties

# finding where in the data frame there is an 'na'
# https://www.geeksforgeeks.org/find-columns-and-rows-with-na-in-r-dataframe/
## funding_data_no_CtyColl = select(funding_data, -c("Collection of Counties"))
funding_data = select(funding_data, -c("Collection of Counties"))

## which(is.na(funding_data_no_CtyColl), arr.ind=T)
# above returns 0 rows, ie
# we find that only the column "Collection of Counties" has 'na'
# we will leave this for now since it may just be a colloquial reference
# unimportant for our data analysis.
# no replacement for na with 0 will be done on this data frame.



# the Costs column has human data, eg $50,890,315.00
# and we need to strip the dollar sign, the commas, 
# and convert them to numbers.  
# ref: https://stackoverflow.com/questions/31944103/convert-currency-with-commas-into-numeric
# we create a new column for this called "Numeric_Cost", 
# but could have potentially done an in-place replacement
funding_data = funding_data %>% 
  mutate(Numeric_Cost = as.numeric(
    gsub( '[$,]', '', funding_data[["Total Costs of OSHPD Projects"]] ) 
  ))

funding_data = funding_data %>%
  mutate( county_name = str_sub( County, 6 )) %>%
  subset(select = -c(County)) %>%
  rename(County = county_name)

# subset HCAI funding data for those "In Closure" state

funding_data_closure <- funding_data %>%
  filter(`OSHPD Project Status` == "In Closure") %>%
  filter(`Data Generation Date` == as_date("2022-08-11")  )

  ## 2022-08-11 the latest avail data

    #group_by(County, Numeric_Cost, `OSHPD Project Status`) %>%  
  ## I (Tin) can no longer get the the gruop_by  to work... 



```



```{r spot checks, echo=FALSE, include=FALSE}


funding_data_closure %>% arrange( Numeric_Cost ) %>% head( 10 )

#View(funding_data)
#View(funding_data_closure)

funding_data %>% select( `Data Generation Date` ) %>% unique() %>% tally() ## 231 rows

selected_counties = c(
  "Inyo",
  "Siskiyou",
  "Mariposa",
  "Modoc",
  "Plumas"
)

funding_data_selected_counties = funding_data %>% 
  #filter( County == "Alameda" ) %>%
  #filter( County == "Inyo" ) %>%
  #filter( County %in% selected_counties  ) %>%
  filter( `OSHPD Project Status` == "In Closure" |
            `OSHPD Project Status` == "In Construction"    ) %>%
  #filter( `OSHPD Project Status` == "Pending Construction"  ) %>%
  #filter( `OSHPD Project Status` == "In Review"  ) %>%
  filter( `Data Generation Date` == 
            as_date( "2022-08-11" ) 
            #as_date( "2020-01-01" ) 
            #as_date( "2016-01-01" ) 
          )

chk1 = 
  ggplot( data = funding_data_selected_counties,
          aes( x = `Data Generation Date`, 
               y = Numeric_Cost             ) ) +
  geom_bar( stat="identity", position="dodge" ) +
  labs( title = 'Prj cost over time' ) + 
  facet_wrap( ~ County )


chk1


## some number may linger in closure for longer than other?
## so taking average of closure amount over say 5 months may lead to skewed data
## really just make sense to just use the last data point.


funding_data_allSources = funding_data %>% 
  #filter( County == "Alameda" ) %>%
  #filter( County == "Mariposa" ) %>%
  #filter( County == "Inyo" ) %>%
  filter( County %in% selected_counties  ) %>%
  filter( `Data Generation Date` > 
            #as_date( "2022-01-01" ) 
            #as_date( "2020-01-01" ) 
            as_date( "2016-01-01" ) 
          )


chk_allSrc = 
  ggplot( data = funding_data_allSources,
          aes( x = `Data Generation Date`, 
               y = Numeric_Cost             ) ) +
  geom_bar( stat="identity", position="dodge" ) +
  labs( title = '1 county, prj stage' ) + 
  facet_wrap( ~ `OSHPD Project Status` )


chk_allSrc


## there is likely more to the story about these costs
## expect to see approved project to have a peak in Review
## then the peak move to Pending Construction, In Construction
## then In Closure
## but maybe their cost is finance book remaining budget account?
## data doesn't fully make sense
## but it is just homework, take a snapshot of "In Closure", and let it be.

```


```{r massage demographics data, echo=FALSE, include=FALSE}


# Load demographics data

demographics_path = 'data/ca_county_demographics.csv'
demographics_data = read_csv( demographics_path, 
                              na = c("", "NA", "-"), 
                              show_col_types=F )

# the first column contains id number, but it is unamed, so renaming it
# rest of the columns have reasonable names, so left them as is.
demographics_data = rename( demographics_data, id="...1")

# View(demographics_data)


# massaging demographics data

demographics_data <- demographics_data %>%
  select(id, name, pop2012, pop12_sqmi, med_age, owner_occ, renter_occ) %>%
  rename( `County` = name )


## Determining Renters vs Homeowners Ratio
demographics_data = demographics_data %>%
  mutate( rent_own_ratio = renter_occ / owner_occ )


demographics_data <- 
  mutate(demographics_data, 
         rural_class=if_else( pop12_sqmi < 20, "rural", "not rural", missing=NULL))

# column to show whether county was rural or not
rural_not_rural <- demographics_data %>% 
  select( County, rural_class, pop12_sqmi, rent_own_ratio, med_age )  

rural_counties = rural_not_rural   %>%
  filter( rural_class == "rural" ) 
  #rename( `Class`  = rural_class )


#### consider a "low pop density" rather than a strict "rural"
#### --> not really good, 49 counties in this grouping

avg_pop_density = mean( demographics_data$pop12_sqmi )
demographics_data = demographics_data %>% 
  mutate( low_pop = if_else( pop12_sqmi < avg_pop_density , 
                                    TRUE,
                                    FALSE,
                                    NULL   ))


```


```{r demographics data complete set high rental, old age, echo=FALSE, include=FALSE}

high_rental_ratio = 
  rural_counties %>% arrange( desc( rent_own_ratio  ) ) 

high_med_age = 
  rural_counties %>% arrange( desc( med_age  ) ) 

## Determine any counties satisfy triple whammy rule

avg_rent_ratio = mean( demographics_data$rent_own_ratio )
avg_age        = mean( demographics_data$med_age  )

demographics_data = demographics_data %>% 
  mutate( high_med_age = if_else( med_age > avg_age , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) %>%
  mutate( high_rental = if_else( rent_own_ratio > avg_rent_ratio , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) 
  
triple_whammy = demographics_data %>%
  filter( rural_class  == "rural" &
          high_rental  == TRUE    &  
          high_med_age == TRUE       )

### so 0 county satisfy the triple whammy rule
### need to modify  selection criteria
```


```{r demographics data RURAL set only, high rental, old age, echo=FALSE, include=FALSE}

## Use only RURAL counties to calculate their averages, and what's above average

## Determine any RURAL counties satisfy double whammy rule
#View(rural_counties)

avg_rent_ratio_rural = mean( rural_counties$rent_own_ratio )
avg_age_rural        = mean( rural_counties$med_age  )



rural_demographics_data = demographics_data %>% 
  mutate( high_med_age_rural = if_else( med_age > avg_age_rural , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) %>%
  mutate( high_rental_rural = if_else( rent_own_ratio > avg_rent_ratio_rural , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) 

#-- c( avg_rent_ratio, avg_rent_ratio_rural, avg_age, avg_age_rural )
#-- sn50 =  rural_demographics_data %>% filter( rural_class == "rural" ) 
```

```{r tmp-double-whammy, include=FALSE, echo=FALSE}

rural_double_whammy = rural_demographics_data %>%
  filter( 
          high_rental_rural  == TRUE    &  
          high_med_age_rural == TRUE       )



rural_double_whammy  # only 3 counties in this list: Lake, Siskiyou, Inyo

rural_demographics_data %>% arrange( rent_own_ratio, med_age ) %>% head( 6 )

```


```{r dem relaxed criteria: old age, high rental in rural standard, echo=FALSE, include=FALSE}

## determine old age as > avg age for full data set
## then pick rural countries, 
## high rental in such rural counties.


#View(rural_counties)


focus_demographics_data = demographics_data %>% 
  mutate( high_rental_rural = if_else( rent_own_ratio > avg_rent_ratio_rural , 
                                    TRUE,
                                    FALSE,
                                    NULL   )) %>%
  mutate( high_med_age = if_else( med_age > avg_age , 
                                    TRUE,
                                    FALSE,
                                    NULL   ))

#-- c( avg_rent_ratio, avg_rent_ratio_rural, avg_age, avg_age_rural )
#-- sn50 =  focus_demographics_data %>% filter( rural_class == "rural" ) 

```

```{r tmp-focus,  include=FALSE, echo=FALSE }

focus_dem = rural_demographics_data %>%
  filter( 
          high_rental_rural  == TRUE    &  
          high_med_age == TRUE       )    %>%
  select( County, rural_class, high_med_age, high_rental_rural )


rural_demographics_data %>%
  select( County, rural_class, med_age, high_med_age, high_rental_rural ) %>%
  arrange( high_rental_rural, desc(med_age) )

#View(focus_dem)  # now have 13 counties to choose from, but not all rural

```



```{r discussion-use-only,  include=FALSE, echo=FALSE}

## decide if need to output any of these tables
## likely need to refine further 

high_rental_ratio %>% head( 6 )

high_med_age %>% head( 6 )

rural_double_whammy %>% head( 6 )

focus_dem %>% head( 13 )

#It doesn't seems like we have high rental (ie > avg_rent) and old age (>avg_age)
#so we will have to judge subjectively
#-Tin


```



```{r calc chronic mortality rate, echo=F, include=F}
# calc chronic mortality rate using census pop12 as denominator

# First need to Join Chronics mortality data (by county, last 5 years)
# and demographics data

demographics_chronic = left_join( demographics_data, 
                                  chronic_by_county, 
                                  by = "County"      ) %>% 
  mutate( pct = ( sum_count / pop2012 ) * 100 )

# pct = "prevalence" percentage
```



```{r join tables, echo=FALSE, include=FALSE}


fund_dem_chron = inner_join( demographics_chronic,
                             funding_data_closure, 
                             by = "County" ) %>%
  mutate( fund_per_cap = Numeric_Cost / pop2012 )

viz_fund_dem_chron = fund_dem_chron %>%
  select( County, 
          pop12_sqmi,
          rural_class,  
          #low_pop,
          med_age,
          high_med_age,
          rent_own_ratio,
          high_rental,
          pct,                    # prevalence,         
          Numeric_Cost, 
          fund_per_cap,
          `Number of OSHPD Projects`,
          )

```


```{r draft-viz-for-discussion-only, echo=FALSE, include=FALSE }

#### for viz, ponder about using these, or derivative thereof 

feel_for_data = viz_fund_dem_chron %>% 
  filter( rural_class == "rural" ) 
  #filter( low_pop == TRUE )         # no good, 49 counties in this grouping

feel_for_data %>% 
  arrange( desc( fund_per_cap ) ) %>% head( 10 )
### above are all 0... which implies no fundings for them... should give them some love!?



```

\newpage 

# Final Report


## Problem Statement (1-2 paragraphs)


++ Aubrey  will write this?

## Methods, for each data source

Source
Years and/or dates of data
Description of cleaning and creating new variables
Analytic methods

++ please review that they make sense, and that i didn't omit anything and met the rubric requirements ++

We used 3 datasets for this project:

 * CA demographics 
 * CA mortality surveillance
 * HCAI healthcare construction funding
 
The first dataset is from the 2012 Census and contains demographics info for each of the 58 counties of California.  It includes info such as population per square mile, median age, number of households who are renters vs owners, ethnicity, genres, etc.   We calculated renter to owner ratio for each county.  Then calculated average age and population density for the whole state and visually inspected the data to see how each county stack up.  We ended up using the National Rural Development Partnership’s definition to determine if a given county's population density is to be classified as rural, for which there were 11.  

The second dataset is the mortality surveillance obtained from the CA Open Data Portal.  It contains a breakdown of total mortality for each county by 15 disease areas.  We used the CDC definition to filter for the chronic diseases, for which 10 fit in this criteria.  The data range from 2014 to 2020, but we were tasked to focus on the last 5 years, thus we applied a filter with Year >= 2016.  As tasked, we also performed filters with Geography_Type == "Occurrence"  and used Strata == "Total Population" to avoid overcounting.  Any missing data were replaced with 0.  Once the data is cleaned, we summed all the disease occurrences within each county.  We joined this with the demographics data to obtain a mortality rate of chronic conditions over 5 years for each county.

The third dataset is the HCAI funding, also obtained from the CA Open Data Portal.  It contains healthcare spending for each county in 4 stages of project progression, updated about every 2 weeks.  We focused on the latest available data, which was Aug 11, 2022, and those with state of "In Closure".  Many rural counties showed up with $0 amount, and we went back to double check our selection code.  It checked out, much fundings are in large populous counties such as LA and Alameda.  It was not that rural counties had no funding, there were fundings for example in the "In Construction" phase, but we decided to focus on "In Closure" to help our improvement plan to drive new spending for rural counties with high and variable mortality rates.

After cleaning and filtering the 3 dataset above, we joined them by county name, whereby we can see which counties has high renters, high chronic mortality rates, and the funding they receive.  

## Results 

++ please improve on this ++


(should this go here?  or in Discussion?)
We tried to see if there were any counties that satisfy all 3 criterias of:

 * Rural by National Rural Development Parthnership's definition.
 * High rental rate (higher than state-wide average)
 * High median age (higher than state-wide average)

But this resulted in 0 county.  As such, our selection criteria were relaxed the selection by removing the 

  rental rate > state average.

From this, we got a list with 8 counties.  We ranked these by renter:owner ratio, and obtained our list of 5 counties to focus: Inyo, Siskiyou, Mariposa, Modoc and Plumas.

When we plotted Mortality Rate for each of the 11 rural counties, we see what our 5 counties selection have the highest rate.  

++ please expand ++

we observe that this list of 8 counties, all had a zero dollar amount for the HCAI funding at the "In Closure" state.

Below ... 

++ please expand ++



## Rural Counties Deserving More HCAI Funding


++ TIN FIX TBD ... i changed it a bit here from Milestone 4, now show 8 counties, 
will find way to highlight them or something ++

Table 1 shows that the 5 counties we selected as deserving more HCAI funding.  
They are the counties of Siskiyou, Inyo, Mariposa, Plumas and Modoc.
All these 5 counties have low population density (they qualify as Rural per National Rural Development Partnership's definition), 
have a high median age, and fairly high ratio of renters.
More importantly, they have one of the highest percentage of chronic diseases mortality,
and no HCAI fundings that are "In Closure" as of 2022-08-11.

As background reference, across all 58 CA counties, we found these statistics:

 * average population density: 665 person per square mile.
 * average rent:own ratio is: 64.7%
 * average age is: 38.5

Note that mortality rate is calculated based on the latest available  population data: 2012.


```{r print quality table, echo=F, include=T }

viz_focus = viz_fund_dem_chron %>%
  filter( rural_class      == "rural",
          high_med_age     == TRUE 
          ##high_rental    == TRUE  
          )

focus_table = viz_focus %>%
  mutate( USD_amount = currency( Numeric_Cost, digits=0L)) %>%
  select( County, 
          pop12_sqmi,
          med_age,
          rent_own_ratio,
          pct,                    # prevalence,         
          USD_amount, 
  ) %>% 
  arrange( desc( pct )) %>%
  rename(
    `Pop Density`       = pop12_sqmi,
    `Median Age`        = med_age,
    `Rent:Own Ratio`    = rent_own_ratio,
    `% Chronic`         = pct,
    `HCAI Fund in 2022` = USD_amount
  ) #%>%    head( 5 ) 

# DT wk 11 reader pdf p6
datatable( 
           focus_table,
           options = list(
                            order=list(5,'desc'),
                            columnDefs=list(
                              list(className='dt-center',targets=1:3)
                            ),
                            dom = 't'  # only want the table, nothing else  
            ),
            caption = "Rural Counties with high median age, rental ratio, and chronic disease rate",
            rownames=FALSE,
          ) %>%
  formatRound( 2:3, 1 ) %>%
  formatRound( 4:5, 2 ) %>%
  formatCurrency( 6, currency="$", digits=0 ) %>%
  formatStyle(1,color="blue") %>%
  formatStyle(2:4,backgroundColor = styleInterval(c(500),c('white','yellow')))

```

## Mortality Rate by County

Below is a graph of Mortality Rate for Chronic diseases (as defined by CDC)
across 11 rural counties (as defined by National Rural Development Partnership)
The 5 counties of focus have the highest mortality rates in this group.

Note that we don't have disease data for Alpine or Sierra county.  

```{r viz: chronic mortality rates by counties, echo=F, include=T}

chronic_focus_counties = inner_join(
    demographics_chronic,
    rural_counties,
    by = "County"                   ) %>%
  mutate( ctyColor = case_when(
    County == "Siskiyou" ~ "red", 
    County == "Inyo"     ~ "darkorange" , 
    County == "Mariposa" ~ "blue", 
    County == "Plumas"   ~ "darkblue",
    County == "Modoc"    ~ "darkcyan",
    TRUE                 ~ "rgb(187, 216, 228)"
  ) )


fig3 = plot_ly( data = chronic_focus_counties ) %>%
    add_trace(
            x = ~County,
            y = ~pct,
            name = 'Chronic Mortality Rates by Counties',
            marker = list(color = ~ctyColor),
            hoverinfo = "text",
            text = ~paste(round(pct, 1), "%" ),  
            type = 'bar') %>%
   layout(
            title = "Chronic Mortality Rates For California's Rural Counties", 
            yaxis = list( title="% Mortality Rate"),
            xaxis = list( title="County", categoryorder='total descending' )
         )

fig3

```
## Chronic disease mortality rates vs HCAI funding

The following boxplot summarizes chronic disease mortality rates from all 
CA counties, grouped according to HCAI funding amounts for in closure projects 
as of August 2022. The funding amounts were categorized as "high" if they were 
above the mean amount, low if they were below the mean, and "no funding" if no 
funding for in closure projects was reported. 
 
(Please describe how this graph help with our narrative on the 5 counties selected)

```{r, echo=F, include=T}

funding_chronic <- funding_data %>% 
  filter( `OSHPD Project Status` == "In Closure") %>%
  filter( `Data Generation Date` == as_date( "2022-08-11")) %>%
  mutate(funding_amount = case_when(
    Numeric_Cost > 12239849 ~ "High Funding", 
    Numeric_Cost == 0 ~ "No Funding", 
    Numeric_Cost < 12239849 ~ "Low Funding"
  )) %>%
  inner_join(demographics_chronic, funding_data_all_counties, by = "County") %>%
  select(pct, County, funding_amount, rural_class, Numeric_Cost)

plot_ly(
  funding_chronic,
  y=~pct,
  color= ~funding_amount,
  type="box"
) %>%
   layout(
    title="Chronic Disease Mortality Rates & HCAI Funding",
    yaxis=list(title="Chronic Disease Rate"))
```

## Table showing most common disease by rural county

This table shows the most common chronic disease in each of the rural counties, while also showing the number of people who have the illness in the year 2020. The counties of Alpine and Sierra did not have chronic disease data available.  As we expand funding in the 5 select counties of focus, special emphasis should be placed on Heart Disease, as that's the most common chronic illness causing high mortality.
*HTD= Heart Disease, CAN= Cancer

```{r, echo=F, include=T}
table_data_e<- inner_join(rural_not_rural, chronic_mortality_data, by= "County") %>% 
select(c("County", "rural_class", "Cause","Count", "Year")) %>% filter(Year%in%2020) %>% 
filter(rural_class=="rural") %>% 
group_by(County,Cause) %>% summarize(count_cause=sum(Count)) %>% arrange(County,desc(c(count_cause))) %>% slice(c(1,11,21,31,41,51,61,71,81,91,101))

table_data_e[1,2]<-"Not Available"
table_data_e[9,2]<-"Not Available"

common_chronic<- kable(table_data_e, col.names=c("County", "Chronic Disease", "Number Reported"),
digits=0, booktabs=T, escape=F, align="ccc", caption="Most Common Chronic Disease by Rural County in 2020")
common_chronic
```

This maybe useful, was left out in Milestone 4.

```{r viz: funding for select counties, include=F, echo=T }

# Visuals saved for possible use in Milestone 5 

# Funding for select counties


counties_funding_comp = funding_data %>%
  filter( `OSHPD Project Status` == "In Closure"  ) %>%
  filter( `Data Generation Date` ==
            as_date( "2022-08-11" )
            #as_date( "2020-01-01" )
            #as_date( "2016-01-01" )
          ) %>%
  arrange( Numeric_Cost )

#fig2 = plot_ly( data = funding_data_1county ) %>%
#fig2 = plot_ly( data = funding_data_selected_counties ) %>%
fig2 = plot_ly( data = counties_funding_comp ) %>%
  add_trace(x = ~`County`,
            y = ~Numeric_Cost,
            type = 'bar',
            name = 'Funding across 58 counties',
            marker = list(color = 'rgb(187, 216, 228)'),
            hoverinfo = "text",
            text = ~paste(round(Numeric_Cost, 0), ' US$')  ) %>%
   layout(
            title="Funding across 58 counties",
            yaxis = list( title="Funding"),
            xaxis = list( title="County", categoryorder='total descending' )
            
         )

fig2


```


## Discussion (interpretation of results)

Maximum of 1,000 words

++ Who will write this?